---
title: "yield loss module outputs and plots"
author: "Jackie"
date: "2024-03-27"
output: html_document
---

```{r setup with library, include=FALSE}
knitr::opts_chunk$set(include = FALSE)
library(tidyverse)
library(readxl)
library(dplyr)
library(ggplot2)
library(plotly)
library(ggpubr)
library(DT)

library(lubridate)
library(patchwork)
#install.packages("patchwork")
```

## Cost of weed project

## Yield loss module - residual weeds in crop

This document is for:
1. checking the model outputs
2. creating template for reporting with plots

```{r import the model output data, message=FALSE, warning=FALSE, include=FALSE}

data_source <- "W:/Economic impact of weeds round 2/model/"
df <- read_excel(paste0(data_source, "yield loss AEZ.xlsx"), sheet = "model cals residual weeds")


df$`Weed free yield t/ha`<- as.double(df$`Weed free yield t/ha`)
```

## For each AEZ and represenative farm (using grossed up values)

1. for all the farms across all the AEZ how much yield is lost in tonnage
2. what is this as a percentage of total yield 

```{r weed 1 and 2 rep farm grossed up, echo=FALSE, message=FALSE, warning=FALSE}

df_GU <- df %>% 
  mutate(
   crop_area_ha_GU  =                  `Crop area ha` * `Gross up factor for AEZ and Crop`,
   area_of_Weed_Infestation_Ha_GU  =   `Area of Weed Infestation (Ha)` * `Gross up factor for AEZ and Crop`,
   yield_loss_tonnes_GU  =             `Yield loss tonnes` * `Gross up factor for AEZ and Crop`,
   revenue_loss_GU  =                  `Revenue loss` * `Gross up factor for AEZ and Crop`
  )

step1_Summary_weed1_2_yld_loss_GU <- df_GU %>% 
  group_by(AEZ_code, `AEZ for report`, Region, Crop) %>% 
  summarise(weed_free_yld_t_ha = mean(`Weed free yield t/ha`, na.rm = TRUE),
            Mean_crop_area_ha_GU =       mean(crop_area_ha_GU , na.rm = TRUE),
            
            sum_area_of_Weed_Infestation_Ha_GU = sum(area_of_Weed_Infestation_Ha_GU, na.rm=TRUE),
            sum_yield_loss_tonnes_GU =           sum(yield_loss_tonnes_GU, na.rm=TRUE),
            sum_revenue_loss_GU =               sum(revenue_loss_GU, na.rm=TRUE)) %>% 
  
  mutate(weed_free_yld_per_farm_tonnes_GU = weed_free_yld_t_ha * Mean_crop_area_ha_GU ) %>%  
  mutate(precenatge_yld_loss_tonnes_per_farm_GU = (sum_yield_loss_tonnes_GU/ weed_free_yld_per_farm_tonnes_GU)*100) 

step1_Summary_weed1_2_yld_loss_GU <- ungroup(step1_Summary_weed1_2_yld_loss_GU)





```



```{r weed 1 and 2 rep farm grossed up summary AEZ, echo=FALSE, message=FALSE, warning=FALSE}

# summaries the data combining AEZ and crop GU
Summary_weed1_2_yld_loss_all_GU <- step1_Summary_weed1_2_yld_loss_GU %>% 
  group_by() %>% 
  summarise(
    sum_yld_on_farm_per_AEZ_GU =  sum(weed_free_yld_per_farm_tonnes_GU, na.rm = TRUE),
    sum_yld_loss_tonnes_AEZ_GU = sum(sum_yield_loss_tonnes_GU, na.rm=TRUE),
    sum_area_weeds_per_AEZ_GU = sum(sum_area_of_Weed_Infestation_Ha_GU, na.rm=TRUE),
    sum_rev_loss_dollars_AEZ_GU = sum(sum_revenue_loss_GU, na.rm=TRUE)) %>% 
  mutate(precenatge_yld_loss_tonnes_per_AEZ_GU = (sum_yld_loss_tonnes_AEZ_GU/ sum_yld_on_farm_per_AEZ_GU)*100) 

Summary_weed1_2_yld_loss_all_GU <- ungroup(Summary_weed1_2_yld_loss_all_GU)

#round percentage
# Summary_weed1_2_yld_loss_all_GU$Summary_weed1_2_yld_loss_all_GU <- 
#   round(Summary_weed1_2_yld_loss_all_GU$Summary_weed1_2_yld_loss_all_GU, 2)
# Summary_weed1_2_yld_loss_all_GU



```


```{r plot - all aez crops yield loss as % production GU, echo=FALSE, message=FALSE, warning=FALSE}
#str(step1_Summary_weed1_2_yld_loss_GU)
#str(Summary_weed1_2_yld_loss_all_GU)
ave_perc_yld_loss_all_GU <-Summary_weed1_2_yld_loss_all_GU$precenatge_yld_loss_tonnes_per_AEZ_GU

#unique(step1_Summary_weed1_2_yld_loss_GU$`AEZ for report`)


step1_Summary_weed1_2_yld_loss_GU %>% 
  ggplot(aes(x= `AEZ for report`, y = precenatge_yld_loss_tonnes_per_farm_GU))+
  geom_point(aes(colour= Crop))+
  geom_boxplot(fill = "white", alpha =0.1)+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.5))+
  labs(
    title = "Yield loss due to residual weeds in crops",
    caption = "Values ARE grossed up. Red line is national average",
    x = "",
    y = "Yield loss as % of tonnes of yield", 
  )+
   geom_hline(yintercept=ave_perc_yld_loss_all_GU, linetype="dashed", 
              color = "red", linewidth=.5)+
  facet_wrap(.~ Region, scales = "free_x")


```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
