---
title: "yield loss module outputs and plots"
author: "Jackie"
date: "2024-03-27"
output: html_document
---

```{r setup with library, include=FALSE}
knitr::opts_chunk$set(include = FALSE)
library(tidyverse)
library(readxl)
library(dplyr)
library(ggplot2)
library(plotly)
library(ggpubr)
library(DT)

library(lubridate)

```

## Cost of weed project

## Yield loss module - residual weeds in crop

This document is for:
1. checking the model outputs
2. creating template for reporting with plots

```{r import the model output data, message=FALSE, warning=FALSE, include=FALSE}

data_source <- "W:/Economic impact of weeds round 2/model/"
df <- read_excel(paste0(data_source, "yield loss AEZ.xlsx"), sheet = "model cals residual weeds")


```

## For each AEZ and represenative farm (not using grossed up values)

1. for all the farms across all the AEZ how much yield is lost in tonnage
2. what is this as a percentage of total yield 

```{r total yield loss all AEZ all crop rep farm not grossed up, echo=FALSE}




```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
